---
title: 'Descriptive Statistics'
subtitle: 'Univariate Statistics'
date: "`r Sys.Date()`"
author: "Your Name"
output:
  html_document: 
    theme: cerulean
    highlight: textmate
    fontsize: 8pt
    toc: yes
    code_download: yes
    toc_float:
      collapsed: no
    df_print: default
    toc_depth: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(qwraps2_markup = "markdown")
library(qwraps2)
library(arsenal)
library(e1071)
library(haven)
library(papeR)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(summarytools)
library(classInt)
library(pastecs)
library(reporttools)
library(desctable)
library(psych)
library(frequency)
library(ggpubr)
library(ggforce)
library(ggdist)
library(gghalves)
library(gtsummary)
library(AER)
download.file("https://github.com/kflisikowski/ds/blob/master/data_apartments.csv?raw=true", destfile ="mieszkania.csv",mode="wb")
apartments <- read.csv("mieszkania.csv",sep=";",dec=",")
```

## Data

In our example this week, we are going to use the fake data - about real
estates in Wroclaw - prices by districts, size of apartments and many
more.

### Preprocessing

As you can see, not all formats of our variables are adapted. We need to
prepare appropriate formats of our variables according to their
measurement scale and future application.

```{r wrangling, include=TRUE}
apartments$district<-as.factor(apartments$district)
apartments$building_type<-as.factor(apartments$building_type)
apartments$rooms<-factor(apartments$rooms,ordered=TRUE)
attach(apartments)
apartments$price_PLN<-as.numeric(apartments$price_PLN)
apartments$price_EUR<-as.numeric(apartments$price_EUR)
```

## Frequency Tables

In the first step of our analysis, we will group our data into a simple
frequency table.

First, let's look at the distribution of housing prices in our sample
and verify tabular validity using the TAI measure:

```{r table, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
etykiety<-c("350-450 kPLN","450-550 kPLN","550-650 kPLN","650-750 kPLN","750-850 kPLN","850-950 kPLN","950-1050 kPLN","1050-1150 kPLN","1150-1250 kPLN","1250-1350 kPLN")
limits<-cut(apartments$price_PLN,seq(350000,1350000,by=100000),labels=etykiety)
tabela1<-freq(limits,type="html")
```

Ok, it looks quite ugly, so let's wrap it up using the 'kable' package:

```{r tai, echo=FALSE}
kbl(tabela1,caption = "Apartments in Wroclaw - prices in kPLN") %>%
    kable_material(c("striped", "hover"))
tab1<-classIntervals(apartments$price_PLN,n=10,style="fixed",fixedBreaks=seq(350000,1350000,by=100000))
jenks.tests(tab1)
```

As we can see - the TAI index is quite high. 0.85 means that we can
accept the proposed construction of the frequency table.

## Basic plots

In this section, we should represent our data using basic (pre-installed
in R) graphics. Select the most appropriate graphs depending on the
scale of the selected variables. Explore the heterogeneity of the
distribution by presenting the data by group (e.g., by neighborhood,
building type, etc.). Don't forget about main titles, labels and
legends. Read more about graphical parameters
[here](http://www.sthda.com/english/wiki/graphical-parameters).

```{r histogram, echo=FALSE}
hist(price_PLN, breaks="FD", col="green", probability = TRUE,
     main="Prices in PLN - Wroclaw")
lines(density(price_PLN[district=="Krzyki"]),col=2)
lines(density(price_PLN[district=="Biskupin"]),col=3)
lines(density(price_PLN[district=="Srodmiescie"]),col=4)
legend("topright", legend=c("Krzyki", "Biskupin", "Srodmiescie"),
       col=c(2,3,4), lty=1:2, horiz=FALSE, box.lty=0, cex=0.8)

```

Note that the `echo = FALSE` parameter has been added to the code
snippet to prevent printing the R code that generated the graph.

```{r boxplot, echo=FALSE}
boxplot(price_PLN~district)
```

## ggplot2 plots

Now, let's use the ***ggplot2*** and ***ggpubr*** libraries to plot.

```{r histogram2, echo=FALSE}
# Density plot of "price_PLN"
#::::::::::::::::::::::::::::::::::::::
density.p <- ggdensity(apartments, x = "price_PLN", 
                       fill = "district", palette = "jco")+
  stat_overlay_normal_density(color = "red", linetype = "dashed")

# Draw the summary table of price_PLN
#::::::::::::::::::::::::::::::::::::::
# Compute descriptive statistics by groups
stable <- desc_statby(apartments, measure.var = "price_PLN",
                      grps = "district")
stable <- stable[, c("district", "length", "mean", "sd")]
# Summary table plot, medium orange theme
stable.p <- ggtexttable(stable, rows = NULL, 
                        theme = ttheme("mOrange"))
# Draw text
#::::::::::::::::::::::::::::::::::::::
text <- paste("Price per apartment by 3 districts - Wroclaw.",
              "Random sample of 200 apartments.",
               sep = " ")
text.p <- ggparagraph(text = text, face = "italic", size = 11, color = "black")
# Arrange the plots on the same page
ggarrange(density.p, stable.p, text.p, 
          ncol = 1, nrow = 3,
          heights = c(1, 0.5, 0.3))
```

Ggplot2 allows you to show the average value for each group using the
**stat_summary()** function. You no longer need to calculate average
values before creating a graph!

```{r boxplot2, echo=FALSE}
ggplot(apartments, aes(x=district, y=price_PLN)) +
    geom_boxplot(alpha=0.7) +
    stat_summary(fun="mean", geom="point", shape=20, size=5, color="red", fill="red") +
 geom_jitter() +
    facet_grid(~building_type) +
    scale_fill_brewer(palette="Set1")

```

### RainCloud Plot

```{r echo=FALSE, message=FALSE, warning=FALSE}
apartments %>% 
  filter(rooms %in% c(1, 2, 3, 4)) %>% 
  ggplot(aes(x = factor(rooms), y = price_PLN, fill = factor(rooms))) +
  
  # add half-violin from {ggdist} package
  stat_halfeye(
    # adjust bandwidth
    adjust = 0.5,
    # move to the right
    justification = -0.2,
    # remove the slub interval
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = 0.12,
    # removing outliers
    outlier.color = NA,
    alpha = 0.5
  ) +
  stat_dots(
    # ploting on left side
    side = "left",
    # adjusting position
    justification = 1.1,
    # adjust grouping (binning) of observations
    binwidth = 0.25
  ) +
# Themes and Labels
  labs(
    title = "RainCloud Plot",
    x = "No. of rooms",
    y = "Prices in PLN",
    fill = "rooms"
  ) +
  coord_flip()
```

### Faceting

Faceting generates small multiples, each showing a different subset of
the data. They are a powerful tool for exploratory data analysis: you
can quickly compare patterns in different parts of the data and see if
they are the same or different. Read more
[here](https://ggplot2-book.org/facet.html).

```{r facet1, echo=FALSE}
plot1 <- ggplot(apartments, aes(price_PLN, rooms)) + 
  geom_abline() +
  geom_jitter(width = 0.1, height = 0.1) 
plot1 + facet_wrap(~district)
```

## Univariate Statistics

Before automatically reporting the full summary table of descriptive
statistics, this time your goal is to measure the central tendency of
the price distribution. Compare the mean, median, and mode along with
positional measures - quantiles - by district and building type or
number of rooms in the apartment.

```{r}
    mean(price_PLN)
    median(price_PLN)
    sd(price_PLN) #standard deviation
    var(price_PLN) #variance
    coeff_var<-sd(price_PLN)/mean(price_PLN) #coefficient of variability %
    coeff_var
    IQR(price_PLN)# difference between quartiles =Q3-Q1 
    sx<-IQR(price_PLN)/2  #interquartile deviation
    coeff_varx<-sx/median(price_PLN) #IQR coefficient of variability %
    coeff_varx
    min(price_PLN)
    max(price_PLN)
    quantile(price_PLN,probs=c(0,0.1,0.25,0.5,0.75,0.95,1),na.rm=TRUE)
```

Ok, we have calculated all of the basic summary statistics above. Let's
wrap them up together now.

```{r kable_report, echo=FALSE}
apartments_list <- split(apartments$price_PLN, apartments$rooms)
inline_plot <- data.frame(rooms = c(1, 2, 3, 4), boxplot = "", histogram = "",
                          line1 = "", line2 = "", points1 = "")
inline_plot %>%
  kbl(booktabs = TRUE) %>%
  kable_paper(full_width = FALSE) %>%
  column_spec(2, image = spec_boxplot(apartments_list)) %>%
  column_spec(3, image = spec_hist(apartments_list)) %>%
  column_spec(4, image = spec_plot(apartments_list, same_lim = TRUE)) %>%
  column_spec(5, image = spec_plot(apartments_list, same_lim = FALSE)) %>%
  column_spec(6, image = spec_plot(apartments_list, type = "p"))

```

### Summary tables

Ok, now we will finally summarize the basic measures of central tendency
for prices by district/building type using the '***kable***' package.
Feel free to customize your final report. See some hints
[here](https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html).

```{r kable_report2, echo=FALSE, message=FALSE, warning=FALSE}
raport <-
  list("Price in PLN" =
       list("Min"       = ~ min(price_PLN),
            "Max"       = ~ max(price_PLN),
            "Q1"        = ~ quantile(price_PLN,0.25),
            "Median" = ~ round(median(price_PLN),2),
            "Q3"        = ~ quantile(price_PLN,0.75),
            "Mean" = ~ round(mean(price_PLN),2),
            "Sd" = ~ round(sd(price_PLN),2),
             "IQR" = ~ round(iqr(price_PLN),2),
            "Sx" = ~ round(iqr(price_PLN)/2,2),
            "Var %" = ~ round((sd(price_PLN)/mean(price_PLN)),2),
            "IQR Var %" = ~ round((iqr(price_PLN)/median(price_PLN)),2),
            "Skewness" = ~  round(skew(price_PLN),2),
             "Kurtosis" = ~  round(kurtosi(price_PLN),2)
            ))
tabela<-summary_table(apartments, summaries = raport, by = c("rooms"))

kbl(tabela,  digits = 2,
  caption="Table 1. Apartments in Wroclaw - prices in PLN by number of rooms.",  col.names = c('1 room', '2 rooms', '3 rooms', '4 rooms')) %>% kable_classic(full_width = F, html_font = "Cambria")%>% kable_styling(bootstrap_options = c("striped", "hover"))
```

### gtsummary

We can calculate easily descriptive statistics also using gtsummary
package:

```{r}
apartments %>%
  select(price_PLN,rooms) %>%
  tbl_summary(label= price_PLN ~ "Price",digits=c(price_PLN)~2,by=rooms,type = all_continuous() ~ "continuous2", statistic = all_continuous() ~ c("{N_nonmiss}", "{median} ({p25}, {p75})", "{min}, {max}"),missing = "no")
```

### dfSummary

dfSummary() creates a summary table with statistics, frequencies and
graphs for all variables in a data frame. The information displayed is
type-specific (character, factor, numeric, date) and also varies
according to the number of distinct values.

When using dfSummary() in R Markdown documents, it is generally a good
idea to exclude a column or two to avoid margin overflow. Since the
Valid and Missing columns are redundant, we can drop either one of them.

```{r warning=FALSE, results="asis"}
dfSummary(apartments,
          plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")
```

To produce optimal results, summarytools has its own version of the base
by() function. It's called stby(), and we use it exactly as we would
by():

```{r results="asis", warning=FALSE}
(stats_by_rooms <- stby(data      = apartments, INDICES   = apartments$rooms, FUN       = descr, stats     = "common", transpose = TRUE))
```

### Tidy Tables

When generating freq() or descr() tables, it is possible to turn the
results into "tidy" tables with the use of the tb() function (think of
tb as a diminutive for tibble). For example:

```{r}
apartments %>%
  descr(stats = "common") %>%
  tb()
```

Here are some examples showing how lists created using stby() or
group_by() can be transformed into tidy tibbles.

```{r}
grouped_descr <- stby(data    = apartments,INDICES = apartments$rooms, FUN     = descr, stats   = "common")

grouped_descr %>% tb()
```

### A Bridge to Other Packages

```{r}
stby(data    = apartments, 
     INDICES = apartments$rooms, 
     FUN     = descr, 
     stats   = "fivenum") %>%
  tb(order = 3) %>%
  kable(format = "html", digits = 2) %>%
  collapse_rows(columns = 1, valign = "top")
```

## Your turn!

Your task this week is to: prepare your own descriptive analysis for the
"CreditCard" dataset (AER package). It is a cross-sectional dataframe on
the credit history for a sample of applicants for a type of credit card.

```{r include=FALSE}
data(CreditCard)
#?CreditCard  read description first
```

Are the yearly incomes (in USD 10,000), credit card expenditures, age,
ratio of monthly credit card expenditure to yearly income -
significantly different for applicants for customers with different
credit risk ("card" variable - factor)?

Prepare a professional data visualizations, descriptive statistics'
tables and interpret them.

## Preprocessing

```{r}
attach(CreditCard)
CreditCard$expenditure<-as.numeric(CreditCard$expenditure)
```

## Frequency Table

```{r my_summary_table}

table<-c("0-300 PLN", "300-600 PLN","600-900 PLN","900-1200 PLN","1200-1500 PLN","1500-1800 PLN","1800-2100 PLN","2100-2400 PLN","2400-2700 PLN","2700-3000 PLN", "3000-3300 PLN")
limits<-cut(CreditCard$expenditure,seq(0,3300,by=300),labels=table)
table1<-freq(limits, type="html")

kbl(table1,caption = "Expenditures") %>%
    kable_material(c("striped", "hover"))
tab1<-classIntervals(CreditCard$expenditure,n=11,style="fixed",fixedBreaks=seq(0,3300,by=300))
jenks.tests(tab1)

```
## Plot

```{r}

# Define the age groups
age_groups <- c("0 - 10", "10 - 20", "20 - 30", "30 - 40", "40 - 50", "50 - 60", "60 - 70", "70 - 80", "80 - 90")

# Create a new column 'age_group' based on the specified ranges
CreditCard$age_group <- cut(CreditCard$age, breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90), labels = age_groups, include.lowest = TRUE)

# Calculate average expenditures for each age group
avg_expenditures <- aggregate(expenditure ~ age_group, data = CreditCard, FUN = mean)

# Bar plot of average expenditures by age group
ggplot(avg_expenditures, aes(x = age_group, y = expenditure)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Average Expenditures by Age Group", x = "Age Group", y = "Average Expenditures") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
    mean(expenditure)
    median(expenditure)
    sd(expenditure) #standard deviation
    var(expenditure) #variance
    coeff_var<-sd(expenditure)/mean(expenditure) #coefficient of variability %
    coeff_var
    IQR(expenditure)# difference between quartiles =Q3-Q1 
    sx<-IQR(expenditure)/2  #interquartile deviation
    coeff_varx<-sx/median(expenditure) #IQR coefficient of variability %
    coeff_varx
    min(expenditure)
    max(expenditure)
    quantile(expenditure,probs=c(0,0.1,0.25,0.5,0.75,0.95,1),na.rm=TRUE)
    
    
    expenditure_list <- split(CreditCard$expenditure, CreditCard$age_group)
inline_plot <- data.frame(age_groups = c("0 - 10", "10 - 20", "20 - 30", "30 - 40", "40 - 50", "50 - 60", "60 - 70", "70 - 80", "80 - 90"), boxplot = "", histogram = "", line1 = "", line2 = "", points1 = "")
inline_plot %>%
  
  kbl(booktabs = TRUE) %>%
  kable_paper(full_width = FALSE) %>%
  column_spec(2, image = spec_boxplot(expenditure_list)) %>%
  column_spec(3, image = spec_hist(expenditure_list)) %>%
  column_spec(4, image = spec_plot(expenditure_list, same_lim = TRUE)) %>%
  column_spec(5, image = spec_plot(expenditure_list, same_lim = FALSE)) %>%
  column_spec(6, image = spec_plot(expenditure_list, type = "p"))

```


```{r}
# Calculate total yearly credit card expenditure
CreditCard$total_expenditure <- CreditCard$expenditure * 12

# Calculate total yearly income
CreditCard$total_income <- CreditCard$income * 1000 * 12

# Calculate ratio of total yearly credit card expenditure to yearly income
CreditCard$ratio <- CreditCard$total_expenditure / CreditCard$total_income

# Group by credit risk (card) and calculate mean ratio
ratio_by_card <- CreditCard %>%
  group_by(card) %>%
  summarise(mean_ratio = mean(ratio))

# Print the result
print(ratio_by_card)
```


From the boxplots, we observe that there are variations in yearly incomes, credit card expenditures, age, and the ratio of monthly credit card expenditure to yearly income across different credit risk categories.

```{r}
# Descriptive analysis and visualization
# Boxplot for yearly incomes by credit risk
ggplot(CreditCard, aes(x = card, y = total_income)) +
  geom_boxplot(fill = "skyblue", color = "blue") +
  labs(x = "Credit Risk", y = "Yearly Income (in USD 10,000)", title = "Yearly Incomes by Credit Risk")

# Boxplot for credit card expenditures by credit risk
ggplot(CreditCard, aes(x = card, y = total_expenditure)) +
  geom_boxplot(fill = "lightgreen", color = "darkgreen") +
  labs(x = "Credit Risk", y = "Credit Card Expenditure", title = "Credit Card Expenditures by Credit Risk")

# Boxplot for age by credit risk
ggplot(CreditCard, aes(x = card, y = age)) +
  geom_boxplot(fill = "lightcoral", color = "darkred") +
  labs(x = "Credit Risk", y = "Age", title = "Age by Credit Risk")

# Boxplot for ratio of monthly credit card expenditure to yearly income by credit risk
ggplot(CreditCard, aes(x = card, y = ratio)) +
  geom_boxplot(fill = "lightyellow", color = "darkorange") +
  labs(x = "Credit Risk", y = "Ratio of Monthly Expenditure to Yearly Income", title = "Ratio by Credit Risk")

```


```{r}
# Descriptive statistics table for yearly incomes by credit risk
yearly_income_summary <- CreditCard %>%
  group_by(card) %>%
  summarise(mean_income = mean(total_income),
            median_income = median(total_income),
            sd_income = sd(total_income),
            min_income = min(total_income),
            max_income = max(total_income))

# Descriptive statistics table for credit card expenditures by credit risk
expenditure_summary <- CreditCard %>%
  group_by(card) %>%
  summarise(mean_expenditure = mean(total_expenditure),
            median_expenditure = median(total_expenditure),
            sd_expenditure = sd(total_expenditure),
            min_expenditure = min(total_expenditure),
            max_expenditure = max(total_expenditure))

# Descriptive statistics table for age by credit risk
age_summary <- CreditCard %>%
  group_by(card) %>%
  summarise(mean_age = mean(age),
            median_age = median(age),
            sd_age = sd(age),
            min_age = min(age),
            max_age = max(age))

# Descriptive statistics table for ratio of monthly credit card expenditure to yearly income by credit risk
ratio_summary <- CreditCard %>%
  group_by(card) %>%
  summarise(mean_ratio = mean(ratio),
            median_ratio = median(ratio),
            sd_ratio = sd(ratio),
            min_ratio = min(ratio),
            max_ratio = max(ratio))

# Print descriptive statistics tables
print("Descriptive Statistics for Yearly Incomes by Credit Risk:")
print(yearly_income_summary)

print("Descriptive Statistics for Credit Card Expenditures by Credit Risk:")
print(expenditure_summary)

print("Descriptive Statistics for Age by Credit Risk:")
print(age_summary)

print("Descriptive Statistics for Ratio of Monthly Expenditure to Yearly Income by Credit Risk:")
print(ratio_summary)
```
